---
title: "Supplementary Material for the article: \nAlternative data sources can fill the gaps in data-poor fisheries"

author: Alexandre M. S. Machado^1,2,3^*, Eduardo L. Hettwer Giehl^4^, Luiza Pacheco Fernandes^1^, Simon N. Ingram^5^, Fábio Gonçalves Daura-Jorge^5^

output:
  bookdown::word_document2:
    fig_caption: yes
    number_sections: yes
    toc: no
    reference_docx: docx_template.docx
---

*^1^ Laboratório de Mamíferos Aquáticos, Universidade Federal de Santa Catarina, CCB/ECZ, Florianópolis, SC 88010-970, Brazil*

*^2^ Programa de Pós-Graduação em Ecologia, Universidade Federal de Santa Catarina, CCB/ECZ, Florianópolis, SC 88010-970, Brazil*

*^3^ Department of Collective Behaviour, Max Planck Institute of Animal Behaviour, Konstanz, Germany*

*^4^ Laboratório de Diversidade e Conservação, Universidade Federal de Santa Catarina, CCB/ECZ, Florianópolis, SC 88010-970, Brazil*

*^5^ School of Biological and Marine Sciences, University of Plymouth, PL48AA, United Kingdom*

`*` Corresponding author: alexandre.marcel@posgrad.ufsc.br

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
# load packages
if(!require(kableExtra)){install.packages("kableExtra")};library(kableExtra)

# set figure resolution
knitr::opts_chunk$set(dpi=600)
```

\newpage

# Bayesian structural time series models

Models, predictions and other data sets presented here are coded in the file `analyses.R`. Check the repository at <https://github.com/machadoams/mullet-alternative-data> for all the R code. 

```{r analyses, cache=FALSE, echo=TRUE, results='hide', fig.keep='none', message=FALSE, warning=FALSE}
# Run all the analyses
source("./R/analyses.R") 

# Eport figure with BSTS results
source("./R/figure_bsts.R") 

# Export figure with cumulative errors
source("./R/figure_bsts_cumulative_errors.R")
```


\newpage

# Self-reporting

## Components of the best model

```{r selfreporting_arx_components, fig.width=6.5, fig.height=4, fig.cap = "Components of state for the self-reporting catch data."}
plot(fitfa.arx.self_reporting, "components")
```

\newpage

## Checking the best model

```{r selfreporting_arx, fig.width=6.5, fig.height=6.5, fig.cap = "Plots of the state contribution to the mean (States, top left panel), the posterior distribution of the residuals (Residuals, top right panel), a time series plot showing the most likely predictor (bottom left panel), and the posterior inclusion probabilities for the predictors of self-reporting catch data (Coefficients, bottom right panel)."}
par(mfrow=c(2,2))
plot(fitfa.arx.self_reporting, main = "States")
plot(fitfa.arx.self_reporting, "residuals", main = "Residuals")
plot(fitfa.arx.self_reporting, "predictors", main = "Predictors")
plot(fitfa.arx.self_reporting, "coefficients", main = "Coefficients")
```

\newpage

## Checking model predictions

```{r tab_pred_sr_arx}
tab_pred_sr_arx <- data.frame(
  year = df_self_reporting[df_self_reporting$year > 2012, 'year'],
  mean = bsts.pred.models$`Self-reporting`$mean,
  median = bsts.pred.models$`Self-reporting`$median,
  ci = paste0(round(bsts.pred.models$`Self-reporting`$interval[1,], 3), "--",
              round(bsts.pred.models$`Self-reporting`$interval[2,], 3)),
  catches = df_self_reporting[df_self_reporting$year > 2012, 'Self-reporting']) %>% 
  dplyr::select('Year' = year,
                'Predicted mean' = mean,
                'Predicted median' = median,
                '95% credible interval' = ci,
                'Yield (tons)' = catches)
```

```{r tabsr}
knitr::kable(tab_pred_sr_arx, 
             format = 'markdown',
             digits = 2, 
             caption = "Predicted mean and median, and 95% credible interval from the best model. Yield shows the empirical data for self-reporting catches from 2013 to 2016.")
```

\newpage

## Checking the posterior distribution

```{r, fig.width=6.5, fig.height=6.5, fig.cap="Full posterior distribution of the selected model using self-reporting catch data. For each year, blue dashed lines show the 95% credible interval, the black dashed line show the predicted mean and red line show the observed catches."}
# get the posterior distribution
distribution_pred_sr_arx <- bsts.pred.models$`Self-reporting`$distribution
colnames(distribution_pred_sr_arx) <- df_self_reporting[df_self_reporting$year > 2012, 'year']

# get mean and CI
dfsr <- data.frame(
  year = df_self_reporting[df_self_reporting$year > 2012, 'year'],
  mean = bsts.pred.models$`Self-reporting`$mean,
  ci.low = round(bsts.pred.models$`Self-reporting`$interval[1,], 3),
  ci.up = round(bsts.pred.models$`Self-reporting`$interval[2,], 3),
  catches = df_self_reporting[df_self_reporting$year > 2012, 'Self-reporting']
)

# reshape and plot posterior distribution by year
reshape2::melt(distribution_pred_sr_arx) %>% 
  mutate(year = as.factor(Var2)) %>% 
  ggplot(aes(x = value, group = Var2)) +
  geom_histogram(binwidth = 0.5) +
  geom_vline(data = dfsr, 
             aes(xintercept = mean),
             linetype = "dashed", size = 1, colour = "black") +
  geom_vline(data = dfsr, 
             aes(xintercept = ci.low),
             linetype = "dashed", size = 1, colour = "blue") +
  geom_vline(data = dfsr, 
             aes(xintercept = ci.up),
             linetype = "dashed", size = 1, colour = "blue") +
  geom_vline(data = dfsr, 
             aes(xintercept = catches),
             linetype = "dashed", size = 1, colour = "firebrick") +
  facet_wrap(~year, ncol = 1) +
  theme_base 
```


\newpage

# Media reports

## Components of the best model

```{r mediareports_components, fig.width=6.5, fig.height=4, fig.cap = "Components of state for the catch data mined from media reports."}
plot(fitfa.arx.media_reports, "components")
```

\newpage

## Checking the best model

```{r mediareports_model, fig.width=6.5, fig.height=6.5, fig.cap = "Plots of the state contribution to the mean (States, top left panel), the posterior distribution of the residuals (Residuals, top right panel), a time series plot showing the most likely predictor (bottom left panel), and the posterior inclusion probabilities for the predictors of catch data mined from media reports (Coefficients, bottom right panel)."}
par(mfrow=c(2,2))
plot(fitfa.arx.media_reports, main = "States")
plot(fitfa.arx.media_reports, "residuals", main = "Residuals")
plot(fitfa.arx.media_reports, "predictors", main = "Predictors")
plot(fitfa.arx.media_reports, "coefficients", main = "Coefficients")
```

\newpage

## Checking model predictions

```{r tab_pred_mr_arx}
tab_pred_mr_arx <- data.frame(
  year = df_media_reports[df_media_reports$year > 2012, 'year'],
  mean = bsts.pred.models$`Media reports`$mean, 
  median = bsts.pred.models$`Media reports`$median,
  ci = paste0(round(bsts.pred.models$`Media reports`$interval[1,], 3),"--", 
              round(bsts.pred.models$`Media reports`$interval[2,], 3)),
  catches = df_media_reports[df_media_reports$year > 2012, 'Media reports']) %>% 
  dplyr::select('Year' = year,
                'Predicted mean' = mean,
                'Predicted median' = median,
                '95% credible interval' = ci,
                'Yield (tons)' = catches)
```

```{r tabmr}
knitr::kable(tab_pred_mr_arx, 
             format = 'markdown',
             digits = 2, 
             caption = "Predicted mean and median, and 95% credible interval from the best model. Yield shows the empirical data mined from media reports from 2013 to 2016.")
```




\newpage

## Checking the posterior distribution

```{r, fig.width=6.5, fig.height=6.5, fig.cap="Full posterior distribution of the selected model using catch data mined from media reports. For each year, blue dashed lines show the 95% credible interval, the black dashed line show the predicted mean and red line show the observed catches."}
# get the posterior distribution
distribution_pred_media_arx <- bsts.pred.models$`Media reports`$distribution
colnames(distribution_pred_media_arx) <- df_media_reports[df_media_reports$year > 2012, 'year']

# get mean and CI
dfmedia <- data.frame(
  year = df_media_reports[df_media_reports$year > 2012, 'year'],
  mean = bsts.pred.models$`Media reports`$mean,
  ci.low = round(bsts.pred.models$`Media reports`$interval[1,], 3),
  ci.up = round(bsts.pred.models$`Media reports`$interval[2,], 3),
  catches = df_media_reports[df_media_reports$year > 2012, 'Media reports']
)

# reshape and plot posterior distribution by year
reshape2::melt(distribution_pred_media_arx) %>% 
  mutate(year = as.factor(Var2)) %>% 
  ggplot(aes(x = value, group = Var2)) +
  geom_histogram(binwidth = 0.5) +
  geom_vline(data = dfmedia, 
             aes(xintercept = mean),
             linetype = "dashed", size = 1, colour = "black") +
  geom_vline(data = dfmedia, 
             aes(xintercept = ci.low),
             linetype = "dashed", size = 1, colour = "blue") +
  geom_vline(data = dfmedia, 
             aes(xintercept = ci.up),
             linetype = "dashed", size = 1, colour = "blue") +
  geom_vline(data = dfmedia, 
             aes(xintercept = catches),
             linetype = "dashed", size = 1, colour = "firebrick") +
  facet_wrap(~year, ncol = 1) +
  theme_base 
```

\newpage

# Beach seiner community

## Components of the best model

```{r beachseiner_components, fig.width=6.5, fig.height=4, fig.cap = "Components of state for the beach seiner community catch data."}
plot(fitfa.arx.beach_seiners, "components")
```

\newpage

## Checking the best model

```{r beachseiner_model, fig.width=6.5, fig.height=6.5,fig.cap = "Plots of the state contribution to the mean (States, top left panel), the posterior distribution of the residuals (Residuals, top right panel), a time series plot showing the most likely predictor (bottom left panel), and the posterior inclusion probabilities for the predictors of a beach seiner community catch data (Coefficients, bottom right panel)."}
par(mfrow = c(2,2))
plot(fitfa.arx.beach_seiners, main = "States")
plot(fitfa.arx.beach_seiners, "residuals", main = "Residuals")
plot(fitfa.arx.beach_seiners, "predictors", main = "Predictors")
plot(fitfa.arx.beach_seiners, "coefficients", main = "Coefficients")
```

\newpage

## Checking model predictions

```{r tab_pred_bs_arx}
tab_pred_bs_arx <- data.frame(
  year = df_beach_seiners[df_beach_seiners$year > 2012, 'year'],
  mean = bsts.pred.models$`Beach seiner community`$mean,
  median = bsts.pred.models$`Beach seiner community`$median,
  ci = paste0(round(bsts.pred.models$`Beach seiner community`$interval[1,], 3), "--", 
              round(bsts.pred.models$`Beach seiner community`$interval[2,], 3)),
  catches = df_beach_seiners[df_beach_seiners$year > 2012, 'Beach seiner community']) %>% 
  dplyr::select('Year' = year,
                'Predicted mean' = mean,
                'Predicted median' = median,
                '95% credible interval' = ci,
                'Yield (tons)' = catches) 
```

```{r tabbs}
knitr::kable(tab_pred_bs_arx, 
             format = 'markdown',
             digits = 2, 
             caption = "Predicted mean and median, and 95% credible interval from the best model. Yield shows the empirical data for the single community beach seine catches from 2013 to 2016.")
```

\newpage

## Checking the posterior distribution

```{r, fig.width=6.5, fig.height=6.5, fig.cap="Full posterior distribution of the selected model using catch data from a beach seiner community. For each year, blue dashed lines show the 95% credible interval, the black dashed line show the predicted mean and red line show the observed catches."}
# get the posterior distribution
distribution_pred_bs_arx <- bsts.pred.models$`Beach seiner community`$distribution
colnames(distribution_pred_bs_arx) <- df_beach_seiners[df_beach_seiners$year > 2012, 'year']

# get mean and CI
dfbs <- data.frame(
  year = df_beach_seiners[df_beach_seiners$year > 2012, 'year'],
  mean = bsts.pred.models$`Beach seiner community`$mean,
  ci.low = round(bsts.pred.models$`Beach seiner community`$interval[1,], 3),
  ci.up = round(bsts.pred.models$`Beach seiner community`$interval[2,], 3),
  catches = df_beach_seiners[df_beach_seiners$year > 2012, 'Beach seiner community']
)

# reshape and plot posterior distribution by year
reshape2::melt(distribution_pred_bs_arx) %>% 
  mutate(year = as.factor(Var2)) %>% 
  ggplot(aes(x = value, group = Var2)) +
  geom_histogram(binwidth = 0.5) +
  geom_vline(data = dfbs, 
             aes(xintercept = mean),
             linetype = "dashed", size = 1, colour = "black") +
  geom_vline(data = dfbs, 
             aes(xintercept = ci.low),
             linetype = "dashed", size = 1, colour = "blue") +
  geom_vline(data = dfbs, 
             aes(xintercept = ci.up),
             linetype = "dashed", size = 1, colour = "blue") +
  geom_vline(data = dfbs, 
             aes(xintercept = catches),
             linetype = "dashed", size = 1, colour = "firebrick") +
  facet_wrap(~year, ncol = 1) +
  theme_base 
```

\newpage

# Session information

```{r echo=FALSE, comment=""}
sessionInfo()
```


```{r table01, include=FALSE}
# Export Table 1 (main text) ----

# ref: https://multithreaded.stitchfix.com/blog/2016/04/21/forget-arima/
PositiveMean <- function(b) {
  b <- b[abs(b) > 0]
  if (length(b) > 0) 
    return(mean(b))
  return(0)
}

fitted.bsts <- list("Self-reporting" = fitfa.arx.self_reporting, 
                    "Media reports" = fitfa.arx.media_reports, 
                    "Beach seiner community" = fitfa.arx.beach_seiners)

bsts.coeff <- data.frame(dataset = NULL,
                         coeff = NULL,
                         inclusion.probs = NULL,
                         var = NULL)

for(i in names(fitted.bsts)){
  tmp.df <- data.frame(row.names = NULL,
    dataset = i,
    coeff = apply(fitted.bsts[[i]]$coefficients[-(1:burnin),], 2, PositiveMean),
    inclusion.probs = colMeans(fitted.bsts[[i]]$coefficients[-(1:burnin),] != 0),
    var = as.character(names(colMeans(fitted.bsts[[i]]$coefficients[-(1:burnin),] != 0))))

  bsts.coeff <- rbind(bsts.coeff, tmp.df)
}

bsts.coeff$coef.probs <- NULL

for(i in 1:nrow(bsts.coeff)){
  bsts.coeff$coef.probs[i] <- paste(round(bsts.coeff$coeff[i], 3), 
                                    paste0("(", round(bsts.coeff$inclusion.probs[i], 3),")",
                                     collapse = ""))
}

reshape2::dcast(bsts.coeff, dataset ~ var, value.var = "coef.probs") %>% 
  dplyr::select("Data set" = 'dataset',
                "Intercept" = '(Intercept)',
                "SST" = 'sst',
                "CPUE" = 'cpue') %>% 
  as.data.frame() %>% 
  write.table(file = "table_01.txt", sep = ",", quote = FALSE, row.names = F)
```

